### 2024-02-14 outputs/paper/paper.qmd
Update paper.qmd

---
title: "The Lasting Shadows of Economic Downturns"
subtitle: "Unveiling the 1980–1982 Recession's Long-Run Effects on Unemployment, Personal Income and Business Pattern in America"
author: "Heyucheng Zhang, Jiwon Choi"
thanks: "Code and data are available at: https://github.com/jwonc4602/Long-Run-Impacts-of-Recession"
date: "12 February 2024"
date-format: "D MMMM YYYY"
abstract: "This study examines the long-term impacts of the 1980–1982 recession, focusing on changes in unemployment rates and personal income in America to assess how economic downturns influence decisions regarding business patterns and human capital investment. Our analysis of historical data reveals that the recession had enduring negative effects on employment and income levels, suggesting that such economic challenges may lead individuals to another business trends as a means to improve their marketability and financial stability. These findings emphasizes the importance of incorporating long-term considerations into economic policy-making, highlighting the need for strategies that mitigate the adverse effects of recessions while fostering investment in human capital."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

# Introduction
*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here](update the link!!!)***

# Data
This paper replicates and extends the analyses from the original study by using a range of datasets obtained from various publicly accessible sources. Stuart has categorized the datasets into distinct files, reflecting their respective roles in the analysis.

## Sources
```{r}
### few rows for unemployment or usa_personal_income data presented here as example ###
```
We used data from two sources for our research. First, we utilized data from the Bureau of Economic Analysis (BEA) Regional Economic Accounts[@BEA]. These datasets were downloaded in March 2015 from [https://apps.bea.gov/regional/downloadzip.cfm](https://apps.bea.gov/regional/downloadzip.cfm) by Stuart. The BEA occasionally revises the methodology used to generate these data, and accordingly, updates all files to reflect these changes. As a result, data downloaded on subsequent dates might present some variations. The datasets of particular interest included 'CA1: Personal Income Summary: Personal Income, Population, Per Capita Personal Income', which covers Personal Income, Population, and Per Capita Personal Income, and 'CA25: Total Full-Time and Part-Time Employment by SIC Industry', offering detailed insights into income components and industry-specific earnings.

Second, we obtained unemployment data for the years 1976-1989 which Stuart obtained from the Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics [@BLS] through email communication with Brian Hannon. This data, covering various counties, is compiled in the file 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' and is available through the Inter-university Consortium for Political and Social Research.

The datasets 'CA1_1969_2013_ALL.csv' and 'CA25_1969_2000_ALL.csv' from the BEA, along with the 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' from the BLS are crucial for analyzing personal income, business patterns, and unemployment trends in our study.

## Methodology and Measurement

# Results
```{r}
### the line graphs of average unemployment rate by four most populated states###
### which is something like simulated plot1###
```

```{r}
### the bar graph change of unemployment for four major states ###
### something like simulated plot2 ###
```

```{r}
### the line graph of personal income per capita by year ###
### something like simulated plot3 ###
```

```{r}
### something like simulated plot4 ###
```

```{r}
### something like simulated plot5 ###
```

# Discussion
## How Recession Impacts on Unemployment
## How Recession Impacts on Personal Income
## How Recession Changes the Business Patterns

## Limitations and Next Steps
Due to the unavailability of datasets, this paper will not explore the further social consequences of economic recessions specifically focusing on how they affect birth rates and education which were initially the main areas of interest. Drawing inspiration from Stuart's study [@10.1257/app.20180055], we quantitatively evaluate the long-term effects of the recession that occurred between 1980 and 1982 on individuals who were in their years—children, adolescents, and young adults—when this economic downturn began. Stuart's findings reveal lasting impacts on educational achievements and income levels. For instance among individuals aged 0–10 in 1979, a decrease of 10 percent in per capita earnings from 1979 to 1982 within their birth county is associated with a decline of 4.7 percentage points in four-year college degree attainment. Additionally, it leads to a drop in earned income by $2,100 as an increase of 1.6 percentage points in the likelihood of experiencing poverty during the period from 2000 to 2013. The negative impact on college graduation rates is particularly notable and consistently significant for individuals aged 0–13 in 1979. Becomes smaller and statistically insignificant for those between 14 and 22 years old. This trend indicates that the recession's influence likely originates from a reduction in childhood capital or sustained financial constraints, among parents affecting their ability to support education. In the future when we have access, to easily accessible datasets researchers will investigate how education and birth rates are impacted.

\newpage
# References